---
title: "Machine Learning Project"
author: "Bel Abarrientos"
date: 
output: 
  html_document:
    keep_md: true
---

## Get Training and Test Data
Some codes were commented to reduce knitr details

```{r getdata}
setwd("D:/$Study/DataScience/ML")
#if(!file.exists("./data")){dir.create("./data")}
#f1url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
setwd("data")
#download.file(f1url,destfile="pml-training.csv")

#f2url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
#download.file(f2url,destfile="pml-testing.csv")

trainAll <- read.csv("pml-training.csv")
testAll <- read.csv("pml-testing.csv")
```

## Load required libraries

```{r loadlibraries}
library(caret)
library(randomForest)
library(e1071)
```

## Explore Data using R Studio
Review of features and what can be considered as predictors, classes
Code commented to reduce details

```{r exploredata}
#head(testAll)
#head(trainAll)
#View(testAll)
#View(trainAll)

```

## Split Training data for model fitting
Derived model will be tested against the remaining training data.  
Further data exploration is done to assess predictors and decide which should be removed.  One review of test data, some columns are completely of NA values, so i decided to removed these columns from the training set.


```{r splitdata}
set.seed(1)
trains1 <- createDataPartition(y=trainAll$classe, p=0.1,list=FALSE)
train1d <- trainAll[trains1,]
train2d <- trainAll[-trains1,]

nzv <- nearZeroVar(train1d, saveMetrics=TRUE)
nzv
```

## Set-up data using Predictor Variables only
NA columns are removed or those with UniquePercent almost zero

```{r removeNAs}
testAllNA <- testAll[,c(12:36, 50:59, 69:83,87:101, 103:112, 125:139, 141:150)]
testAllNoNA <- testAll[,-c(12:36, 50:59, 69:83,87:101, 103:112, 125:139, 141:150)]
colNa <- colnames(testAllNA)

train1dNoNA <- train1d[,!(names(train1d) %in% colNa)]
train2dNoNA <- train2d[,!(names(train2d) %in% colNa)]
testNoNA <- testAll[,!(names(testAll) %in% colNa)]

train1fl <- train1dNoNA[8:60]
train2fl <- train2dNoNA[8:60]
testfl   <-  testNoNA[8:59]
```

## Model Fit using subset of training data and randomForest
various size of predictor variables were tested from 5, 8, 10, 12, 14, 15, 16, 20, 40.  mtry=12 provide the best accuracy result 

At this number of predictors, the result is 95.76% accurate, out of sample error is very minimal 0.11% the most (class B).  Overall dataset out-of-bag error is 5.04%. 



```{r trainpredict, cache=TRUE}
set.seed(2)
trainmod <- randomForest(formula = classe~.,data=train1fl, mtry=12, importance=TRUE)
print(trainmod)

testmod <- predict(trainmod, newdata=train2fl)
confusionMatrix(testmod,train2fl$classe)

testClasse <- predict(trainmod, newdata=testfl)
testClasse
```

## Compare random forests with 10-fold cross validation
Support Vendor Machines (SVM), ten repititions of 10-fold cross-validation using errorest functions in ipred package is compared with randomForest.
Code commented due to lengthy processing.

RandomForest error rate is consistent with SVM error rate using small training model, rf mean 0.5229 , SVM  mean 0.5662.  

```{r crossvaliderrortest, cache=TRUE}
library(ipred)
set.seed(12)
ptm <- proc.time()
error.RF <- numeric(10) 
#for (i in 1:10) {
#    error.RF[i] <- errorest(classe ~., data=train1fl, model=randomForest, mtry=12)$error
#    }
proc.time() - ptm
#summary(error.RF)

set.seed(112)
error.SVM <- numeric(10) 
#for (i in 1:10) {
#    error.SVM[i] <- errorest(classe ~., data=train1fl, model=svm, cost=10, gamma=1.5)$error
#    }
#summary(error.SVM)
```

## Generate the submission files

```{r gensubmitfiles}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(testClasse)

```
Citation:
Data used in this assignment is from Human Activity Recognition by:
Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.
http://groupware.les.inf.puc-rio.br/har